---
title: "Tidy Tuesday Week 5: County Data"
author: "Ted Laderas"
date: "4/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(skimr)
library(cowplot)
county_data <- read.csv("data/acs2015_county_data.csv")
```

## Skim the Variables

https://github.com/rfordatascience/tidytuesday/blob/master/data/acs2015_county_data.csv

```{r}
skimr::skim(county_data)
```

## How many counties does each state have?

```{r}
county_data %>% count(State) %>% arrange(desc(n))
```

## County Data by State

```{r}
per_cap_state <- county_data %>% ggplot(aes(x=reorder(State, IncomePerCap, FUN = median), y=IncomePerCap)) + geom_boxplot() + theme(axis.text.x = element_text(angle=90)) + ggtitle("Income Per Capita By State")

per_cap_state
```

## Let's Look at Rhode Island

Here we look at a subset of the variables and compare the boxplots of the Rhode Island values to the other states.

```{r}
county_data %>% mutate(is_ri = ifelse(State == "Rhode Island", "RI", "OTHER")) %>%
  tidyr::gather("variable", "value", -CensusId, -State, -County, -TotalPop, -Men, -Women, -is_ri, -Employed, -Citizen, -Income, -IncomeErr,-IncomePerCap, -IncomePerCapErr) %>%
  ggplot(aes(x=variable, y=value, group=is_ri, fill=is_ri)) + geom_boxplot() + facet_grid(.~variable,scales = "free_x")  + theme(axis.text.x = element_text(angle=90)) 
```

If we pick out some of the variables that show a large difference, we get the below plot:

```{r}
large_diff_variables <- c("Carpool", "ChildPoverty", "Construction", "Poverty", "PrivateWork", "Production", "Professional")

large_diff <- county_data %>% mutate(is_ri = ifelse(State == "Rhode Island", "RI", "OTHER")) %>%
  tidyr::gather("variable", "value", -CensusId, -State, -County, -TotalPop, -Men, -Women, -is_ri, -Employed, -Citizen, -Income, -IncomeErr,-IncomePerCap, -IncomePerCapErr) %>%
  filter(variable %in% large_diff_variables) %>%
  ggplot(aes(x=variable, y=value, group=is_ri, fill=is_ri)) + geom_boxplot() + facet_grid(.~variable,scales = "free_x")  + theme(axis.text.x = element_text(angle=90), 
  strip.background = element_blank(), strip.text.x = element_blank()) +
  ggtitle("Why is Rhode Island So Rich Compared to Other States?")

large_diff
```

Finally, we plot both per capita income by state and the Rhode Island differences in one single plot.

```{r fig.height=10, fig.width=7}
cowplot::plot_grid(per_cap_state, large_diff, nrow=2)

```